Supporting Information : The importance of a complete description of dependence between variables for ecology and related fields

### 2019-05-10 Paper.Rmd
PCR:09may2019-edits included

\noindent Reid: School of Biological & Marine Sciences, University of Plymouth, Drake Circus, Plymouth PL4 8AA, United Kingdom

@@

All branches of ecology study relationships among environmental and biological variables. However, ubiquitously used approaches to studying such relationships, based on correlation and regression, provide only a small slice of the complex information contained in the relationships. Other statistical tools exist that provide a complete description of relationships between variables, based on the concept of the *copula*; they are applied in finance, neuroscience and other fields, but not in ecology. We here explore the concepts that underpin copulas and examine the potential for those concepts to improve our understanding of ecology. We find that informative copula structure in dependences between variables is exceedingly common across all the environmental, species-trait, phenological, population, community, and ecosystem functioning dataset we considered. Many datasets exhibited asymmetric tail dependence, whereby two variables are more strongly related in their left compared to right tails, or vice versa. Copula structure and tail dependence can help reveal causal relationships between variables. We describe mechanisms by which observed copula structure and tail dependence can arise in ecological data, including a Moran-like effect whereby dependence structures between environmental variables are inherited by ecological variables; and asymmetric or nonlinear influences of environments on ecological variables, such as under Liebig's law of the minimum. We describe consequences of copula structure, including impacts on extinction risk and the stability through time of ecosystem services. By documenting the importance of a complete description of dependence, advancing conceptual frameworks, and demonstrating a powerful approach, we aim to elevate copula reasoning to widespread use in ecology.

@@

We approached our overall goal of exploring the potential of 
copula approaches for ecology by addressing the 
following specific questions. (Q1) Do datasets in ecology 
have copula structure distinct from that of a Gaussian/normal distribution (here called 
non-normal copula structure)? Do they show tail dependence distinct from that 
of a normal distribution, and in particular do they show asymmetric tail 
dependence? (Q2) If so, what are some possible causes/mechanisms of this 
non-normal copula structure? And (Q3) what are the consequences of non-normal 
copula structure and tail dependence for our understanding of ecology and for 
applications? To address Q1, we analyze several datasets, including 
environmental, species-trait, phenological, population, community, and 
ecosystem functioning data. We selected these datasets to span multiple 
sub-fields and organizational levels of ecology. The second section of the 
paper introduces the data, and the third and fourth sections use the data to 
address Q1. The fifth and sixth sections, which address Q2, use simple models to 
explore mechanisms. The seventh and eigth sections, which address Q3, use both
data and models to explore a variety of consequences. The ninth section is the Discussion. 
Copula approaches have been used to great consequence in neuroscience [@onken2009], 
bioinformatics 
[@Kim2008], medical research [@Emura2016], direct study of environmental 
variables [@Serinaldi2008; @She2018; @Goswami2018; @li2013], and finance [@Li2000]. 
To our knowledge, these concepts have never been used in ecology. 
We argue that benefits to ecology of using a complete 
desription of dependence will be substantial.

@@

<!--summary table with info of all data used in Paper-->
```{r tab_summary_data_info, echo=F, results="asis",message=F}
library(kableExtra)
library(dplyr)
tab_data_info<-as.data.frame(matrix(NA,nrow=8,ncol=4))
colnames(tab_data_info)<-c("Data","Category","No. of measurements we used", "References")
tab_data_info$Data<-c("Soil C and N","Bird body masses and BMR", "Mammal body masses and BMR", 
                      "Green spruce and other aphid species abundances", "Leaf-curling plum and other aphid first flight dates","Ceratium furca and other plankton taxa abundances", "Plant diversity and aboveground biomass","Methane flux")
tab_data_info$Category<-c("Environmental","Species-trait","Species-trait",
                          "Population","Phenological","Population","Community level","Ecosystem functioning")
tab_data_info$`No. of measurements we used`<-c(paste(nrow(d_soilCN)," locations",sep=""),
                                               paste(nrow(d_birdsBMR)," birds",sep=""),
                                               paste(nrow(d_mammalsBMR)," mammals",sep=""),
                                               "10 locations with atleast 30 yrs. of timeseries for each 20 sp.",
                                               "10 locations with atleast 30 yrs. of timeseries for each 20 sp.",
                                               "14 locations with atleast 45 yrs. of timeseries for each 22 taxa",
                                               paste(nrow(d_cc[[5]])," plots",sep=""),
                                               "13 locations with atleast 50 dates of data")
tab_data_info$References<-c("US Rapid Carbon Assessment database (RaCA), Wills et al. (2014)", "McNab (2009)","McNab (2008)",
                            "Rothamsted Insect Survey","Rothamsted Insect Survey","Continuous Plankton Recorder Survey",
                            "Cedar Creek Ecosystem Science Reserve, biodiversity experiment e120",
                            "Great Miami Wetland Mitigation Bank, Smyth et al. (2019)")

knitr::kable(tab_data_info, 
             format="latex", align="l",
             caption = "Summary table for the data we used. Bolds entries are the multivariate data, the rest are bivariate datasets (see Data section). Basal metabolic rate=BMR. \\label{tab_data_info}", 
             booktabs = T,linesep = "\\addlinespace") %>%
             column_spec(1:4,width="2.8 cm")%>%
             row_spec(c(4,5,6,8), bold=T)
```

@@

Across essentially all our datasets, copula structures were typically non-normal and 
commonly showed asymmetric tail 
dependence. To make results easier to absorb: we first 
present results in for an example bivariate dataset; then we summarize results
for all bivariate datasets; then we present results for an example multivariate dataset; then we summarize for all multivariate datasets.

@@

We present green spruce aphid abundance results (Table \ref{tab_data_info})
as an example multivariate analysis.
Independence was rejected for each pair of the 10 sampling locations. 
Best-fitting (lowest-AIC) copulas were non-N
for the large majority of location pairs (Table \ref{SM-tab_count_bestcop}), and 
AIC weights for best-fitting copulas (Table \ref{SM-tab_count_AICw}) 
were typically substantially 
larger than N-copula weights (Table \ref{SM-tab_count_AICw_normal}), answering 
the first part of Q1. Goodness of fit tests 
(Tables \ref{SM-tab_count_p_CvM}, \ref{SM-tab_count_p_KS}) in every case 
failed to reject the hypothesis that the AIC-best copula family was also 
an objectively adequate description of 
the data; the collection of 
copula families we used was apparently sufficiently broad that all pairs of 
locations could be characterized by at least one of them.
Model-averaged lower- (Table \ref{SM-tab_count_LTdep_AICw}) and upper-tail
(Table \ref{SM-tab_count_UTdep_AICw}) dependence statistics were exclusively 
positive and were frequently large, thereby differing 
from what an N copula would give (i.e., 0), and answering the second part of 
Q1. These data have greater tail dependence 
(lower and upper) than would be expected from an N-copula null hypothesis. 
Model-averaged lower- minus upper-tail dependence statistics 
(Table \ref{SM-tab_count_LTmUT}) were positive 
for all but a few locations pairs (`r sum((RES_aphid_count$LTdep_AICw-RES_aphid_count$UTdep_AICw)>0,na.rm=T)` 
out of `r (dim(RES_aphid_count$gfc_p_CvM)[1]^2-dim(RES_aphid_count$gfc_p_CvM)[1])-2*RES_aphid_count$num_indep_loc_pair`). Thus the spatial synchrony of rarity in the green spruce aphid is stronger 
than the spatial synchrony of outbreaks. This also answers the third part of Q1.
Nonparametric statistics verified that tail dependence was asymmetric for the green 
spruce aphid abundance data (Appendix \ref{SM-Q1multvarnonparam}, \ref{SM-spatial_resamp}).

<!--***DAN: You could maybe move the next para to sup mat, though that would complicate the
presentation of skewness results below. Only do it if you are desperate to cut.-->
<!--multivariate summary-->
Table \ref{tab_multivar_summary} summarizes Q1 results from
multivariate data. Results supported the 
conclusions that non-N copula structure, non-N tail dependence, and asymmetric tail 
dependence were common. Most locations pairs were non-independent 
(Table \ref{tab_multivar_summary}, row 2). The large majority of non-independent location 
pairs had best-fitting copula that was not the N copula (Table \ref{tab_multivar_summary}, 
row 3), and AIC weights of best-fitting copulas were 
better than those of N copulas (Table \ref{tab_multivar_summary}, rows 4-5).
Best-fitting copulas were nearly always considered an adequate fit 
(Table \ref{tab_multivar_summary}, row 6). Some of the datasets (green spruce aphid
abundance, \textit{Ceratium furca} abundance, methane flux) showed stronger 
or marginally stronger lower- than upper-tail dependence 
(Table \ref{tab_multivar_summary}, rows 7-9), whereas leaf-curling plum aphid first flight
data showed the reverse. Asymmetry results were verified by nonparametric approaches 
<!--(e.g., Table \ref{tab_multivar_summary}, row 10)-->
(Appendix \ref{SM-Q1multvarnonparam}, \ref{SM-spatial_resamp}). Detailed results are in 
Tables \ref{SM-tab_count_bestcop}-\ref{SM-tab_methane_resamp} and 
Figs \ref{SM-fig_aphid_count}-\ref{SM-fig_methane}.
We also carried out the same analyses for abundance and first-flight data for the 18 
aphid species for which we had data other than the green spruce and leaf-curling plum aphids,
as well as for the 21 plankton taxa for which we had abundance data other than the
\textit{Ceratium furca} (results not shown); conclusions supported that 
non-N copula structure and tail dependence, and asymmetry of tail dependence, are 
common.

@@

knitr::kable(tab_multivar_summary[1:9,], 
             format="latex", escape = F,
             caption.short = "Summary of results for multivariate datasets",
             caption = "Summary of results for multivariate datasets. Rows 3-9 of the table were computed for the non-independent pairs (row 2) only. A well-fitting location pair (row 6) was one for which the best-fitting copula had $p$-values $>0.01$ for both the Cramer-von Mises and Kolmogorov-Smirnov goodness of fit tests. \\label{tab_multivar_summary}",
             booktabs=T, linesep = "\\addlinespace",
             col.names = c(" ","Green spruce aphid abundance",
                        "Leaf-curling plum aphid first flight",
                        "\\textit{Ceratium furca} abundance","Methane flux")
             )%>%column_spec(2:5,width="2.5cm")%>%
             column_spec(1, width = "7cm")%>%kable_styling(latex_options = "basic",position = "center")
```

@@

<!--what we did and why is it important to study?-->
We showed that non-N copula structure and asymmetric tail dependence are 
exceedingly common in datasets across multiple sub-disciplines
and levels of organization in ecology, although no published 
research we are aware of has 
taken this structure into account. Via simple models, we hypothesized three 
mechanisms that may cause non-N copula structures. One
mechanism was via a Moran-like effect, whereby dependencies between environmental variables 
beget similar dependencies between ecological variables in different locations.
A second mechanism produces tail dependence via nonlinear environmental influences
on ecological variables. We discuss below how commonly these and other
potential mechanisms may opperate. 
We also demonstrated important consequences of 
non-N and tail-dependent copula structure in ecology. First, the skewness
of a spatial-average time series is influenced by tail dependencies between its 
constituent time series. Thus right-tail-dependent local time series 
lead to "spiky" spatially averaged time series, with large outbreaks.
These observations could have implications for numerous exploited and pest species,
for which the spatial average time series have applied importance.
Second, metapopulation extinction risk can be accentuated or reduced by tail dependence
properties of environmental variables. We also showed effects of copula 
structure on Taylor's law, which is has been applied in many fields.
In our view, our results demonstrate that comprehensive understanding 
of many ecological phenomena is impossible without 
a complete characterization of dependence between variables. 
Copula approaches are
well developed, statistically [@joe2014_dependence; @nelsen2006_copula], and 
computer implementations exist (e.g., the `copula` and 
`VineCopula` packages in R); ecologists can start applying these tools
immediately. We created a variety of interrelated randomization procedures
(Appendices \ref{SM-surrog_test}, \ref{SM-Surrog_multi}, \ref{SM-sect:copsurrognd}) 
that build upon existing methods and may be especially useful for ecological applications.

>>>>

\noindent Reid: School of Biological & Marine Sciences, University of Plymouth, Drake Circus, Plymouth PL4 8AA, UK; Continuous Plankton Recorder Survey, The Marine Biological Association, The Laboratory, Citadel Hill, Plymouth PL1 2PB, UK

@@

All branches of ecology study relationships among environmental and biological variables. However, ubiquitously used approaches to studying such relationships, based on correlation and regression, provide only a small slice of the complex information contained in the relationships. Other statistical tools exist that provide a complete description of relationships between variables, based on the concept of the *copula*; they are applied in finance, neuroscience and other fields, but not in ecology. We here explore the concepts that underpin copulas and examine the potential for those concepts to improve our understanding of ecology. We find that informative copula structure in dependences between variables is exceedingly common across all the environmental, species-trait, phenological, population, community, and ecosystem functioning datasets we considered. Many datasets exhibited asymmetric tail dependence, whereby two variables are more strongly related in their left compared to right tails, or vice versa. Copula structure and tail dependence can help reveal causal relationships between variables. We describe mechanisms by which observed copula structure and tail dependence can arise in ecological data, including a Moran-like effect whereby dependence structures between environmental variables are inherited by ecological variables; and asymmetric or nonlinear influences of environments on ecological variables, such as under Liebig's law of the minimum. We describe consequences of copula structure, including impacts on extinction risk and the stability through time of ecosystem services. By documenting the importance of a complete description of dependence, advancing conceptual frameworks, and demonstrating a powerful approach, we aim to elevate copula reasoning to widespread use in ecology.

@@

We approached our overall goal of exploring the potential of 
copula approaches for ecology by addressing the 
following specific questions. (Q1) Do datasets in ecology 
have copula structure distinct from that of a Gaussian/normal distribution (here called 
non-normal copula structure)? Do they show tail dependence distinct from that 
of a normal distribution, and in particular do they show asymmetric tail 
dependence? (Q2) If so, what are some possible causes/mechanisms of this 
non-normal copula structure? And (Q3) what are the consequences of non-normal 
copula structure and tail dependence for our understanding of ecology and for 
applications? To address Q1, we analyze several datasets, including 
environmental, species-trait, phenological, population, community, and 
ecosystem functioning data. We selected these datasets to span multiple 
sub-fields and organizational levels of ecology. The second section of the 
paper introduces the data, and the third and fourth sections use the data to 
address Q1. The fifth and sixth sections, which address Q2, use simple models to 
explore mechanisms. The seventh and eigth sections, which address Q3, use both
data and models to explore a variety of consequences. The ninth section is the Discussion. 
Copula approaches have been used to great consequence in neuroscience [@onken2009], 
bioinformatics 
[@Kim2008], medical research [@Emura2016], direct study of environmental 
variables [@Serinaldi2008; @She2018; @Goswami2018; @li2013], and finance [@Li2000]. 
To our knowledge, these concepts have never been used in ecology. 
We argue that benefits to ecology of using a complete 
description of dependence will be substantial.

@@

<!--summary table with info of all data used in Paper-->
```{r tab_summary_data_info, echo=F, results="asis",message=F}
library(kableExtra)
library(dplyr)
tab_data_info<-as.data.frame(matrix(NA,nrow=8,ncol=4))
colnames(tab_data_info)<-c("Data","Category","No. of measurements we used", "References")
tab_data_info$Data<-c("Soil C and N","Bird body masses and BMR", "Mammal body masses and BMR", 
                      "Green spruce and other aphid species abundances", "Leaf-curling plum and other aphid first flight dates","\\textit{Ceratium furca} and other plankton taxa abundances", "Plant diversity and aboveground biomass","Methane-flux")
tab_data_info$Category<-c("Environmental","Species-trait","Species-trait",
                          "Population","Phenological","Population","Community level","Ecosystem functioning")
tab_data_info$`No. of measurements we used`<-c(paste(nrow(d_soilCN)," locations",sep=""),
                                               paste(nrow(d_birdsBMR)," birds",sep=""),
                                               paste(nrow(d_mammalsBMR)," mammals",sep=""),
                                               "10 locations with atleast 30 yrs. of timeseries for each 20 sp.",
                                               "10 locations with atleast 30 yrs. of timeseries for each 20 sp.",
                                               "14 locations with atleast 45 yrs. of timeseries for each 22 taxa",
                                               paste(nrow(d_cc[[5]])," plots",sep=""),
                                               "13 locations with atleast 50 dates of data")
tab_data_info$References<-c("US Rapid Carbon Assessment database (RaCA), Wills et al. (2014)", "McNab (2009)","McNab (2008)",
                            "Rothamsted Insect Survey","Rothamsted Insect Survey","Continuous Plankton Recorder Survey",
                            "Cedar Creek Ecosystem Science Reserve, biodiversity experiment e120",
                            "Great Miami Wetland Mitigation Bank, Smyth et al. (2019)")

knitr::kable(tab_data_info, 
             format="latex", align="l",escape=F,
             caption = "Summary table for the data we used. Bold entries are the multivariate data, the rest are bivariate datasets (see Data section). Basal metabolic rate=BMR. \\label{tab_data_info}", 
             booktabs = T,linesep = "\\addlinespace") %>%
             column_spec(1:4,width="2.8 cm")%>%
             row_spec(c(4,5,6,8), bold=T)
```

@@

Across essentially all our datasets, copula structures were typically non-normal and 
commonly showed asymmetric tail 
dependence. To make results easier to absorb: we first 
present results for an example bivariate dataset; then we summarize results
for all bivariate datasets; then we present results for an example multivariate dataset; then we summarize for all multivariate datasets.

@@

We present green spruce aphid abundance results (Table \ref{tab_data_info})
as an example multivariate analysis.
Independence was rejected for each pair of the 10 sampling locations. 
Best-fitting (lowest-AIC) copulas were non-N
for the large majority of location pairs (Table \ref{SM-tab_count_bestcop}), and 
AIC weights for best-fitting copulas (Table \ref{SM-tab_count_AICw}) 
were typically substantially 
larger than N-copula weights (Table \ref{SM-tab_count_AICw_normal}), answering 
the first part of Q1. Goodness of fit tests 
(Tables \ref{SM-tab_count_p_CvM}, \ref{SM-tab_count_p_KS}) in every case 
failed to reject the hypothesis that the AIC-best copula family was also 
an objectively adequate description of 
the data; the collection of 
copula families we used was apparently sufficiently broad that all pairs of 
locations could be characterized by at least one of them.
Model-averaged lower- (Table \ref{SM-tab_count_LTdep_AICw}) and upper-tail
(Table \ref{SM-tab_count_UTdep_AICw}) dependence statistics were exclusively 
positive and were frequently large, thereby differing 
from what an N copula would give (i.e., 0), and answering the second part of 
Q1. These data have greater tail dependence 
(lower and upper) than would be expected from an N-copula null hypothesis. 
Model-averaged lower- minus upper-tail dependence statistics 
(Table \ref{SM-tab_count_LTmUT}) were positive 
for all but a few location pairs (`r sum((RES_aphid_count$LTdep_AICw-RES_aphid_count$UTdep_AICw)>0,na.rm=T)` 
out of `r (dim(RES_aphid_count$gfc_p_CvM)[1]^2-dim(RES_aphid_count$gfc_p_CvM)[1])-2*RES_aphid_count$num_indep_loc_pair`). Thus the spatial synchrony of rarity in the green spruce aphid is stronger 
than the spatial synchrony of outbreaks. This also answers the third part of Q1.
Nonparametric statistics verified that tail dependence was asymmetric for the green 
spruce aphid abundance data (Appendix \ref{SM-Q1multvarnonparam}, \ref{SM-spatial_resamp}).

<!--***DAN: You could maybe move the next para to sup mat, though that would complicate the
presentation of skewness results below. Only do it if you are desperate to cut.-->
<!--multivariate summary-->
Table \ref{tab_multivar_summary} summarizes Q1 results from
multivariate data. Results supported the 
conclusions that non-N copula structure, non-N tail dependence, and asymmetric tail 
dependence were common. Most location pairs were non-independent 
(Table \ref{tab_multivar_summary}, row 2). The large majority of non-independent location 
pairs had best-fitting copula that was not the N copula (Table \ref{tab_multivar_summary}, 
row 3), and AIC weights of best-fitting copulas were 
better than those of N copulas (Table \ref{tab_multivar_summary}, rows 4-5).
Best-fitting copulas were nearly always considered an adequate fit 
(Table \ref{tab_multivar_summary}, row 6). Some of the datasets (green spruce aphid
abundance, \textit{Ceratium furca} abundance, methane-flux) showed stronger 
or marginally stronger lower- than upper-tail dependence 
(Table \ref{tab_multivar_summary}, rows 7-9), whereas leaf-curling plum aphid first flight
data showed the reverse. Asymmetry results were verified by nonparametric approaches 
<!--(e.g., Table \ref{tab_multivar_summary}, row 10)-->
(Appendix \ref{SM-Q1multvarnonparam}, \ref{SM-spatial_resamp}). Detailed results are in 
Tables \ref{SM-tab_count_bestcop}-\ref{SM-tab_methane_resamp} and 
Figs \ref{SM-fig_aphid_count}-\ref{SM-fig_methane}.
We also carried out the same analyses for abundance and first-flight data for the 18 
aphid species for which we had data other than the green spruce and leaf-curling plum aphids,
as well as for the 21 plankton taxa for which we had abundance data other than the
\textit{Ceratium furca} (results not shown); conclusions supported that 
non-N copula structure and tail dependence, and asymmetry of tail dependence, are 
common.

@@

knitr::kable(tab_multivar_summary[1:9,], 
             format="latex", escape = F,
             caption.short = "Summary of results for multivariate datasets",
             caption = "Summary of results for multivariate datasets. Rows 3-9 of the table were computed for the non-independent pairs (row 2) only. A well-fitting location pair (row 6) was one for which the best-fitting copula had $p$-values $>0.01$ for both the Cramer-von Mises and Kolmogorov-Smirnov goodness of fit tests. \\label{tab_multivar_summary}",
             booktabs=T, linesep = "\\addlinespace",
             col.names = c(" ","Green spruce aphid abundance",
                        "Leaf-curling plum aphid first flight",
                        "\\textit{Ceratium furca} abundance","Methane-flux")
             )%>%column_spec(2:5,width="2.5cm")%>%
             column_spec(1, width = "7cm")%>%kable_styling(latex_options = "basic",position = "center")
```

@@

<!--what we did and why is it important to study?-->
We showed that non-N copula structure and asymmetric tail dependence are 
exceedingly common in datasets across multiple sub-disciplines
and levels of organization in ecology, although no published 
research we are aware of has 
taken this structure into account. Via simple models, we hypothesized three 
mechanisms that may cause non-N copula structures. One
mechanism was via a Moran-like effect, whereby dependencies between environmental variables 
beget similar dependencies between ecological variables in different locations.
A second mechanism produces tail dependence via nonlinear environmental influences
on ecological variables. We discuss below how commonly these and other
potential mechanisms may opperate. 
We also demonstrated important consequences of 
non-N and tail-dependent copula structure in ecology. First, the skewness
of a spatial-average time series is influenced by tail dependencies between its 
constituent time series. Thus right-tail-dependent local time series 
lead to "spiky" spatially averaged time series, with large outbreaks.
These observations could have implications for numerous exploited and pest species,
for which the spatial average time series have applied importance.
Second, metapopulation extinction risk can be accentuated or reduced by tail dependence
properties of environmental variables. We also showed effects of copula 
structure on Taylor's law, which has been applied in many fields.
In our view, our results demonstrate that comprehensive understanding 
of many ecological phenomena is impossible without 
a complete characterization of dependence between variables. 
Copula approaches are
well developed, statistically [@joe2014_dependence; @nelsen2006_copula], and 
computer implementations exist (e.g., the `copula` and 
`VineCopula` packages in R); ecologists can start applying these tools
immediately. We created a variety of interrelated randomization procedures
(Appendices \ref{SM-surrog_test}, \ref{SM-Surrog_multi}, \ref{SM-sect:copsurrognd}) 
that build upon existing methods and may be especially useful for ecological applications.

### 2019-05-10 SupportingInformation.Rmd
PCR:09may2019-edits included

\textbf{Ecosystem functioning data on methane flux.} The Great Miami Wetland Mitigation Bank  was at 
$36^{\circ}46^{'}51^{''}$ N, $84^{\circ}20^{'}26^{''}$ W, in Trotwood, Montgomery County, Ohio. Additional information 
on the history of the site and soil types is detailed in @Jarecke2016. Briefly, the closed static flux chamber method 
was used [@Holland1999] to measure $\text{CH}_{4}$ fluxes at 28 locations across a hydrologic gradient (upland to 
seasonal wetland) at daily to weekly interval from September 2015 to September 2016. See @smyth2019using for
additional details on greenhouse gas flux methods for this site. We considered 13 locations for which data were
gathered for at least 50 dates.

@@

<!--Load and format the data : methane flux-->
```{r read_format_methane_data, echo=F}
#latlon_methane_old<-readRDS("./Data/Methane_data/chamberLocations.rds")
#write.csv(latlon_methane_old,"./Data/Methane_data/chamberLocations.csv",row.names = F)
latlon_methane<-read.csv("./Data/Methane_data/chamberLocations.csv")
#identical(latlon_methane_old,latlon_methane)  # checked : they are identical :)

>>>>

\textbf{Ecosystem functioning data on methane-flux.} The Great Miami Wetland Mitigation Bank  was at 
$36^{\circ}46^{'}51^{''}$ N, $84^{\circ}20^{'}26^{''}$ W, in Trotwood, Montgomery County, Ohio. Additional information 
on the history of the site and soil types is detailed in @Jarecke2016. Briefly, the closed static flux chamber method 
was used [@Holland1999] to measure $\text{CH}_{4}$ fluxes at 28 locations across a hydrologic gradient (upland to 
seasonal wetland) at daily to weekly interval from September 2015 to September 2016. See @smyth2019using for
additional details on greenhouse gas flux methods for this site. We considered 13 locations for which data were
gathered for at least 50 dates.

@@

<!--Load and format the data : methane-flux-->
```{r read_format_methane_data, echo=F}
#latlon_methane_old<-readRDS("./Data/Methane_data/chamberLocations.rds")
#write.csv(latlon_methane_old,"./Data/Methane_data/chamberLocations.csv",row.names = F)
latlon_methane<-read.csv("./Data/Methane_data/chamberLocations.csv")
#identical(latlon_methane_old,latlon_methane)  # checked : they are identical :)
